---
title: "Dealing with Errors"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  dev = "ragg_png",
  dpi = 300
)
```

```{r setup}
library(rATTAINS)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(ggtext)
library(mpsTemplates)
library(webmockr)
mpsTemplates::noto_dark_geom_defaults()
webmockr::enable()
```

There are a number of errors that you might encounter using this rATTAINS. Here
is a list of potential errors and fixes. Feel free to raise an issue if I missed
something.


## Network Connectivity

The following error message likely indicates an issue connecting to the EPA
server:



```{r echo=TRUE, eval=FALSE}
state_summary(organization_id = "TCEQMAIN", reporting_cycle = "2022")
```

```{r echo=FALSE}
warning("No connection to <https://attains.epa.gov> available!",
        call. = FALSE)
```

Potential issues/fixes:
 
 - Check your network connection.
 - Check [attains.epa.gov](attains.epa.gov). If you are able to connect, a warning notice about accessing U.S. Government information systems should show in your web browser.
 - Occasionally proxy systems used in corporate IT systems cause issues with connections (see: [https://stackoverflow.com/questions/59796178/r-curlhas-internet-false-even-though-there-are-internet-connection](https://stackoverflow.com/questions/59796178/r-curlhas-internet-false-even-though-there-are-internet-connection)). I've tried to account for this in the package, but you might run into occasional issues.
 

## Server Response

The server might also return http code messages. The most common will be 404 or 429. rATTAINS will generally provide a simple message and error when this is encountered:

```{r echo=FALSE}
webmockr::stub_request("get",
                       "https://attains.epa.gov/attains-public/api/actions?actionIdentifier=R8-ND-2018-03&summarize=N&returnCountOnly=N") |> 
  to_return(status = 429)
```

```{r}
actions(action_id = "R8-ND-2018-03")
```

```{r echo=FALSE, message=FALSE}
webmockr::disable()
```


Potential issues/fixes:

- Wait until the server is responsive.
- Make less frequent requests.

## Parsing Errors

The default behavior in rATTAINS is to parse JSON data downloaded from the API to one or more dataframes. These are returned as a single dataframe or list of dataframes depending on the function. rATTAINS also tries to flatten the data as much as possible. This design choice **might** have been a mistake because it can become a source of errors if the data returned by the API changes or is inconsistent. As of version 1.0.0 of the package the `.unnest` argument was added to most functions. By setting `.unnest=FALSE` many of these problems should be avoided.

Default behavior:

```{r}
state_summary(organization_id = "TDECWR", 
              reporting_cycle = "2016")
```


Using `.unnest=FALSE` returns nested columns. The tidyr family of `unnest()` functions is an easy way to flatten this data:

```{r}
df <- state_summary(organization_id = "TDECWR", 
                    reporting_cycle = "2016",
                    .unnest = FALSE)
df

df |>
  tidyr::unnest(reporting_cycles) |> 
  tidyr::unnest(water_types) |> 
  tidyr::unnest(use_attainments)
```

If the above option doesn't work, rATTAINS can also provide the raw JSON data from the API. The [tibblify](https://cran.r-project.org/package=tibblify) 📦️ and [jsonlite](https://cran.r-project.org/package=jsonlite) 📦 provide tools to convert JSON to nested lists then tibbles pretty easily. First, use the `tidy=FALSE` argument to return the unparsed JSON string, then uses jsonlite to convert that data to a nested list, then tibblify to convert to a nested dataframe!

```{r}
raw_data <- state_summary(organization_id = "TDECWR", 
                    reporting_cycle = "2016",
                    tidy = FALSE)

list_data <- jsonlite::fromJSON(raw_data,
                                simplifyVector = FALSE,
                                simplifyDataFrame = FALSE,
                                flatten = FALSE)

df <- tibblify::tibblify(list_data$data,
                         unspecified = "drop")
df$reportingCycles
```

